---
title: "Transform Data"
output: html_document
---

```{r setup}
library(dplyr)
library(babynames)
library(nycflights13)
library(skimr)
library(ggplot2)
```

## Babynames

```{r}
babynames
skim(babynames)
# View(babynames) # can't Knit a document with View in it
my_skim <- skim_with(numeric = sfl(p25 = NULL, p75=NULL))
```

## Your Turn 1
Run the skim_with() command, and then try skimming `babynames` again to see how the output is different. Look at the examples and try some other variations.
```{r}
my_skim(babynames)
```

## Data types in R

## Your Turn 2

How many dimensions does a vector have?

## Your Turn 3

What happens in your Environment when you run this code?
In your RMarkdown document?

```{r}
vec <- c(1, 2, 3, 10, 100)
vec 
```

## Your Turn 4

Which of these are numbers? What are the others? How can you tell?
1 "1" "one"

## Your Turn 5
Make a vector that contains the number 1, the letter R, and the logical TRUE.

What class of data is the vector?


```{r}
y <- c(1, "R", TRUE)
y
class(c1)
x <- c(1 , 2, 3)
x
class(x)
"x"

```


## Your Turn 6

Another way to see the class of an object is in the Environment pane. Does the Environment agree with what you found using class()?

## Select

```{r}
select(babynames, name, prop)
select(babynames, n)
dplyr::select(.data = babynames, ... = name, prop) # this is the same as above just spelled out
```

## Your Turn 7

Alter the code to select just the `n` column:

```{r}
select(babynames, -prop) # Use negative to drop a column
select(babynames, ends_with("x"))
select(babynames, starts_with("n"))
select(babynames, name, prop)
```


## Consider

Which of these is NOT a way to select the `name` and `n` columns together?

```{r}
select(babynames, -c(year, sex, prop))
select(babynames, name:n)
select(babynames, starts_with("n"))
select(babynames, ends_with("n"))
```

## Filter

```{r}
filter(babynames, name == "Amelia") # find all the amelia's in the data set
filter(babynames, prop >= 0.08)
filter(babynames, name=="Sea")
filter(babynames, sex=="M", year==2000)
filter(babynames, sex == "F" & name == "Sea")
filter(babynames, (n == 5| n == 6)  & year == 1880)
filter(babynames, name == "Acura" | name == "Lexus" | name == "Yugo")
filter(babynames, name %in% c("Acura", "Lexus", "Yugo"))
babynames %>% 
  filter(name %in% c("Acura", "Lexus", "Yugo")) %>% 
  count(name)

```

## Your Turn 8

Show:

* All of the names where prop is greater than or equal to 0.08  
* All of the children named "Sea"  
* All of the names that have a missing value for `n`  

```{r}
filter(babynames, is.na(n))

```

## Your Turn 9

Use Boolean operators to alter the code below to return only the rows that contain:

* Girls named Sea  
* Names that were used by exactly 5 or 6 children in 1880  
* Names that are one of Acura, Lexus, or Yugo

```{r}
filter(babynames, name == "Sea" | name == "Anemone")
```

## Arrange

```{r}
arrange(babynames, n) # default is smalles to largest
```

## Your Turn 10

Arrange babynames by `n`. Add `prop` as a second (tie breaking) variable to arrange on. Can you tell what the smallest value of `n` is?

```{r}
arrange(babynames, n, prop)
```

## desc

```{r}
arrange(babynames, desc(n))
```

## Your Turn 11

Use `desc()` to find the names with the highest prop.
Then, use `desc()` to find the names with the highest n.

```{r}
arrange(babynames, desc(prop))
arrange(babynames, desc(n))
```

## Steps and the pipe

```{r}
babynames %>%
  filter(year == 2015, sex == "M") %>%
  select(name, n) %>%
  arrange(desc(n))
```

## Your Turn 12

Use `%>%` to write a sequence of functions that: 

1. Filter babynames to just the girls that were born in 2015  
2. Select the `name` and `n` columns  
3. Arrange the results so that the most popular names are near the top.

```{r}
babynames %>% 
  filter(sex == "F", year == 2015) %>% 
  select(name, n) %>% 
  arrange(desc(n))

```

## Your Turn 13

1. Trim `babynames` to just the rows that contain your `name` and your `sex`  
2. Trim the result to just the columns that will appear in your graph (not strictly necessary, but useful practice)  
3. Plot the results as a line graph with `year` on the x axis and `prop` on the y axis

```{r}
babynames %>% 
  filter(name == "Gabriel", sex == "M") %>% 
  select(year, prop) %>% 
  ggplot(mapping = aes(x = year, y = prop)) +
  geom_line()

babynames %>% 
  filter(name == "Joey", sex == "M") %>% 
  select(year, prop) %>% 
  ggplot() +
  geom_line(mapping = aes(x = year, y = prop))

```

## Your Turn 14

Use summarise() to compute three statistics about the data:

1. The first (minimum) year in the dataset  
2. The last (maximum) year in the dataset  
3. The total number of children represented in the data

```{r}
babynames %>% 
  summarise(
    total = sum(n), # variable called total
    max = max(n)    # variable called max
  )

babynames %>% 
  summarise(
    first = min(year),
    last = max(year),
    total = sum(n)
  )

```

## Your Turn 15

Extract the rows where `name == "Khaleesi"`. Then use `summarise()` and a summary functions to find:

1. The total number of children named Khaleesi
2. The first year Khaleesi appeared in the data

```{r}
babynames %>% 
  filter(name == "Khaleesi") %>% 
  summarise(
    total = sum(n),
    first = min(year)
  )

babynames %>% 
  filter(name == "Khaleesi") %>% 
  select(year, prop) %>% 
  ggplot() +
  geom_line(mapping = aes(x = year, y = prop))

babynames %>% 
  summarise(
    n = n(),
    nname = n_distinct(name)
  )

babynames %>% 
  group_by(sex) %>% 
  summarise(
    total = sum(n)
  )

```

## Toy data for transforming

```{r}
# Toy dataset to use
pollution <- tribble(
       ~city,   ~size, ~amount, 
  "New York", "large",       23,
  "New York", "small",       14,
    "London", "large",       22,
    "London", "small",       16,
   "Beijing", "large",      121,
   "Beijing", "small",       56
)
```

## Summarize

```{r}
pollution %>% 
 summarise(mean = mean(amount), sum = sum(amount), n = n())

pollution %>% 
  summarise(
    mean = mean(amount), 
    sum = sum(amount),
    n = n()
  )
```

```{r}
pollution %>% 
  group_by(city) %>%
  summarise(mean = mean(amount), sum = sum(amount), n = n())

pollution %>% 
  group_by(city) %>% 
  summarise(
    mean = mean(amount),
    sum = sum(amount),
    n = n()
  )

pollution %>% 
  group_by(city, size) %>% 
  summarise(
    mean = mean(amount),
    sum = sum(amount),
    n = n()
  )
```


## Your Turn 16

Use `group_by()`, `summarise()`, and `arrange()` to display the ten most popular baby names. Compute popularity as the total number of children of a single gender given a name.

```{r}

babynames %>% 
  group_by(sex, name) %>% 
  summarise(
    total = sum(n)
    ) %>% 
    arrange(desc(total)) %>% 
   head(10) %>% # selecting first six rows 
  ggplot() +
  geom_col(aes(y = total, x = name, fill = sex))

babynames %>% 
  group_by(name, sex) %>% 
  summarise(
    total = sum(n)
    ) %>% 
  summarise(
    n = n(), 
    total = sum(total)
    ) # changes the grouping

babynames %>% 
  mutate(
    percent = round(prop*100, 2) # 2 refers to number of decimals
  )


```

## Your Turn 17

Use grouping to calculate and then plot the number of children born each year over time.

```{r}

```

## Ungroup

```{r}
babynames %>%
  group_by(name, sex) %>% 
  summarise(total = sum(n)) %>% 
  arrange(desc(total))
```

## Mutate

```{r}
babynames %>%
  mutate(percent = round(prop*100, 2))
```

## Your Turn 18

Use `min_rank()` and `mutate()` to rank each row in `babynames` from largest `n` to lowest `n`.

```{r}
babynames %>% 
  mutate(
    rank = min_rank(desc(n))
  )

babynames %>% 
  mutate(
    rank = min_rank(desc(prop))
  )

babynames %>% 
  group_by(year, sex) %>% 
  mutate(
    rank = min_rank(desc(prop))
  )

babynames %>% 
  group_by(year, sex) %>% 
  mutate(
    rank = min_rank(desc(prop))
  ) %>% 
  group_by(name, sex) %>% 
  summarise(
    score = median(rank)
  ) %>% 
  arrange(score)

```

## Your Turn 19

Compute each name's rank _within its year and sex_. 
Then compute the median rank _for each combination of name and sex_, and arrange the results from highest median rank to lowest.

```{r}

```

## Joining data

```{r}
library(nycflights13)
```

## Flights data
```{r}
flights
skim(flights)
```

## Toy data

```{r}
band <- tribble(
   ~name,     ~band,
  "Mick",  "Stones",
  "John", "Beatles",
  "Paul", "Beatles"
)

instrument <- tribble(
    ~name,   ~plays,
   "John", "guitar",
   "Paul",   "bass",
  "Keith", "guitar"
)

instrument2 <- tribble(
    ~artist,   ~plays,
   "John", "guitar",
   "Paul",   "bass",
  "Keith", "guitar"
)
```

## Mutating joins


## Types of joins

```{r}
band %>% left_join(instrument, by = "name")
band %>% right_join(instrument, by = "name")
band %>% full_join(instrument, by = "name")
band %>% inner_join(instrument, by = "name")
```


## Your Turn 20

Which airlines had the largest arrival delays?  Work in groups to complete the code below.

1. Join `airlines` to `flights`
2. Compute and order the average arrival delays by airline. Display full names, no codes.


```{r}
flights %>%
  drop_na(arr_delay) %>%
     #something!     %>%
  group_by(   #something!  ) %>%
     #something!      %>%
  arrange(  #something! )       
```

## Different names

```{r}
band %>% left_join(instrument2, by = c("name" = "artist"))
```

```{r}
airports %>% left_join(flights, by = c("faa" = "dest"))
```

## Your Turn 21

Find the average delay by destination airport. Find the average delay by _origin_ airport.

```{r}
__________ %>%
  drop_na(arr_delay) %>%
  group_by(__________) %>% 
  summarise(delay = mean(arr_delay)) %>%
  _________(_________, by = ___________)
```

## Filtering joins


```{r}
band %>% semi_join(instrument, by = "name")
band %>% anti_join(instrument, by = "name")
```

## Your Turn 22

How many airports in airports are serviced by flights originating in New York (i.e. flights in our dataset?) 
Notice that the column to join on is named faa in the airports dataset and dest in the flights dataset.

```{r}

```

***

# Take aways

* Extract variables with `select()`  
* Extract cases with `filter()`  
* Arrange cases, with `arrange()`  

* Make tables of summaries with `summarise()`  
* Make new variables, with `mutate()`  
* Do groupwise operations with `group_by()`

* Connect operations with `%>%`  

* Use `left_join()`, `right_join()`, `full_join()`, or `inner_join()` to join datasets
* Use `semi_join()` or `anti_join()` to filter datasets against each other


<!-- This file by Amelia McNamara is licensed under a Creative Commons Attribution 4.0 International License, adapted from the orignal work at https://github.com/rstudio/master-the-tidyverse by RStudio. -->
