---
title: "Data Types"
output: html_document
---

```{r setup}
library(tidyverse)
library(lubridate)
library(fivethirtyeight)
library(forcats)
library(stringr)
library(ggplot2)
library(hms)
```

## Warm-up / Review

Using the data `gss_cat`, find the average hours of tv watched (`tvhours`) for each category of marital status (`marital`).

```{r}
gss_cat
```

## Your Turn 1

What kind of object is the `marital` variable?  

```{r}
gss_cat
```

Brainstorm with your neighbor, all the things you know about that kind of object.

# Factors

```{r}
# Example of a factor
eyes <- factor(x = c("blue", "green", "green"), 
               levels = c("blue", "brown", "green"))
eyes
unclass(eyes)

```

```{r}
x <- factor(c("characters", "in", "the", "streets"))
y <- factor(c("integers", "in", "the", "sheets"))

c(x, y)

```


```{r}

str(gss_cat$race)
levels(gss_cat$race)
nlevels(gss_cat$race)
class(gss_cat$race)

forcats::fct_count(gss_cat$race)

gss_cat$race %>%  # reverse coded 
  fct_infreq() %>% 
  fct_rev() %>% 
  levels()

gss_cat %>% 
  count(race)

gss_cat %>% 
  pull(marital) %>% 
  levels()

gss_cat %>% 
  filter(!is.na(race)) %>% 
  mutate(new_race = fct_rev(race)) %>% 
  count(new_race, sort = TRUE)

gss_cat %>% 
  filter(!is.na(race)) %>% 
  mutate(new_race = fct_lump(race)) %>% 
  count(new_race, sort = TRUE)

gss_cat %>% 
  filter(!is.na(race)) %>% 
  mutate(new_race = fct_lump(race, n = 3)) %>% # can specify the number of categories to lump 
  count(new_race, sort = TRUE)

```


## Your Turn 2

Fix your summary of average hours of tv watched (`tvhours`) by marital status (`marital`), to drop missing values in `tvhours`, then create a plot to examine the results.

```{r}
# two ways of removing missing values 
gss_cat %>%
  drop_na(tvhours) %>% 
  group_by(marital) %>% 
  summarise(avg_tvhours = mean(tvhours)) 

gss_cat %>%
  group_by(marital) %>% 
  summarise(avg_tvhours = mean(tvhours, na.rm = TRUE)) 

```
## Your Turn 3

Fill in the blanks (`   `) to explore the average hours of tv watched by religion.

```{r, error = TRUE}
gss_cat %>%
  drop_na(tvhours) %>%
  group_by(relig) %>%
  summarise(
    avg_tvhour = mean(tvhours)
  ) %>%
  ggplot() +
    geom_point(mapping = aes(x = relig, y = avg_tvhour))

```

## Consider

Why is this plot not very useful?

```{r}
gss_cat %>%
  drop_na(tvhours) %>%
  group_by(denom) %>%
  summarise(avg_tvhours = mean(tvhours)) %>%
  ggplot() +
    geom_point(mapping = aes(x = avg_tvhours,
      y = fct_reorder(denom, avg_tvhours)))

gss_cat %>%
  drop_na(tvhours) %>%
  group_by(relig) %>%
  summarise(avg_tvhours = mean(tvhours)) %>%
  ggplot() +
    geom_point(mapping = aes(x = avg_tvhours,
      y = fct_reorder(relig, avg_tvhours))) # fct_order takes one vector  

gss_cat %>%
  drop_na(tvhours) %>%
  group_by(relig, age) %>%
  summarise(avg_tvhours = mean(tvhours)) %>%
  ggplot() +
    geom_point(mapping = aes(x = avg_tvhours,
      y = fct_reorder2(relig, age, avg_tvhours), color = age)) # fct_reorder2 is used when you have 2 vectors to plot
```
## Your Turn 4

Edit the code to also relabel some other Baptist denominations:

* "Baptist-dk which" -> "Baptist - Don't Know"    
* "Other baptists" -> "Baptist = Other"

```{r}
gss_cat %>% 
  mutate(
    denom = fct_recode(denom,"Baptist - Southern" = "Southern baptist"), # rename categories within a variable
    denom = fct_recode(denom, "Baptists - Dont Know" = "Baptist-dk which"),
    denom = fct_recode(denom, "Baptists - Other" = "Other baptists"),
  ) %>%
  pull(denom) %>%
  levels() # givres you the categories of a variable 

gss_cat %>% # this is another way just using "," 
  mutate(
    denom = fct_recode(denom,"Baptist - Southern"   = "Southern baptist", 
                             "Baptists - Dont Know" = "Baptist-dk which", 
                             "Baptists - Other"     = "Other baptists")
  ) %>%
  pull(denom) %>%
  levels() # gives you the categories of a variable 

```
## Your Turn 5

What does the function `detect_denom()` do?

```{r}
detect_denom <- function(x){
  case_when(
    str_detect(x, "[Bb]ap")  ~ "Baptist", # uses string detect to rename 
    str_detect(x, "[Pp]res") ~ "Presbyterian",
    str_detect(x, "[Ll]uth") ~ "Lutheran",
    str_detect(x, "[Mm]eth") ~ "Methodist",
    TRUE ~ x
  )
}

gss_cat %>% pull(denom) %>% levels() %>% detect_denom()
```

```{r}
gss_cat %>%
 drop_na(tvhours) %>%
 mutate(
   denom_higher = denom %>%
     fct_relabel(detect_denom) %>%
     fct_reorder(tvhours, mean)
 ) %>%
  group_by(denom_higher, denom) %>%
 summarise(
   avg_tvhours = mean(tvhours)
   ) %>%
 ggplot() +
 geom_point(mapping = aes(x = avg_tvhours, y = denom_higher))

```



# Strings

With your neighbor, predict what these might return:

```{r}
strings <- c("Apple", "Pineapple", "Orange")

str_detect(strings, pattern = "pp")
str_detect(strings, pattern =  "apple")
str_detect(strings, pattern = "[Aa]pple")
```

Then run them!

# Times and Dates

```{r}
# An example for times/dates
births <- US_births_1994_2003 %>%
  select(date, births)
```

## Your Turn 7

For each of the following formats (of the same date), pick the right `ymd()` function to parse them:

```{r}
ymd("2018 Feb 01")
mdy("2-1-18")
dmy("01/02/2018")
ymd_hms("2017-01-11 01:30:55")
wday(ymd("2018-02-01"))
wday(ymd("2018-02-01"), label = TRUE)
wday(ymd("2018-02-01"), label = TRUE, abbr = FALSE)


```

## Your Turn 8

Fill in the blanks to:

* Extract the month from date. 
* Extract the year from date.
* Find the total births for each year/month.
* Plot the results as a line chart.

```{r, error = TRUE}
births %>%
  mutate(
    year = year(date),
    month = month(date, label = TRUE)
    ) %>%
  group_by(year, month) %>%
  summarise(
    total_births = sum(births)) %>%
  ggplot() + 
    geom_line(aes(x = month, y = total_births, group = year, color = year))

```

```{r}
hms::hms(seconds = 56, min = 34, hour = 12)
unclass(hms(56, 34, 12))
```


# Take Aways

Dplyr gives you three _general_ functions for manipulating data: `mutate()`, `summarise()`, and `group_by()`. Augment these with functions from the packages below, which focus on specific types of data.

Package   | Data Type
--------- | --------
forcats   | factors
stringr   | strings
hms       | times
lubridate | dates and times

<!-- This file by Charlotte Wickham is licensed under a Creative Commons Attribution 4.0 International License, adapted from the orignal work at https://github.com/rstudio/master-the-tidyverse by RStudio. -->
