---
title: "Iteration"
output: html_document
---

```{r setup}
library(tidyverse)

# Toy data
set.seed(1000)
exams <- list(
  student1 = round(runif(10, 50, 100)),
  student2 = round(runif(10, 50, 100)),
  student3 = round(runif(10, 50, 100)),
  student4 = round(runif(10, 50, 100)),
  student5 = round(runif(10, 50, 100))
)

extra_credit <- list(0, 0, 10, 10, 15)
```

## Your Turn 1

What kind of object is `mod`?  Why are models stored as this kind of object?

```{r}
mod <- lm(price ~ carat + cut + color + clarity, data = diamonds)
View(mod)
```

## Consider

What's the difference between a list and an **atomic** vector?

Atomic vectors are: "logical", "integer", "numeric" (synonym "double"), "complex", "character" and "raw" vectors.

## Your Turn 2

Here is a list:

```{r}
a_list <- list(nums = c(8, 9), 
            log = TRUE,    
            cha = c("a", "b", "c"))
```

Here are two subsetting commands. Do they return the same values? Run the code chunk above, _and then_ run the code chunks below to confirm

```{r}
a_list["nums"] # train and contents
a_list["log"] 
a_list["cha"] 
```

```{r}
a_list$nums # contents but no train
a_list$log
a_list$cha
a_list[["nums"]] 
a_list[["log"]] 
a_list[["cha"]] 
```

## Your Turn 3

What will each of these return? Run the code chunks to confirm.

```{r}
vec <- c(-2, -1, 0, 1, 2)
abs(vec)
```

```{r, error = TRUE}
lst <- list(-2, -1, 0, 1, 2)
abs(lst) # cannot take the absolute value of a list 
lst
view(lst)
```

## Your Turn 4

Run the code in the chunks. What does it return?

```{r}
list(student1 = mean(exams$student1),
     student2 = mean(exams$student2),
     student3 = mean(exams$student3),
     student4 = mean(exams$student4),
     student5 = mean(exams$student5))
```

```{r}
library(purrr)
map(exams, mean) # use purrr to exam lists - you can write a for loop but we use purrr now 

exams %>% 
  map(mean)
```

## Your Turn 5

Calculate the variance (`var()`) of each student’s exam grades.

```{r}
exams %>% 
  map(var) # get the variance of each students exam scores 

exams %>% # other options are map_chr() "character vector", map_dbl() "double vector", map_int() "integer vector", map_lgl() "logical vector", map_df() "data frame"
  map_dbl(var)

exams %>%
map_dbl(quantile, prob = 0.9)

exams %>%
map(length) %>%
map_lgl(all.equal, 10) # 10 corresponds to the number of entries for each element - here it is student exam scores 

```

## Your Turn 6

Calculate the max grade (`max()`)for each student. Return the result as a vector.

```{r}
exams %>% 
  map_dbl(max) # locate the maximum exam score for each student 
```

## Your Turn 7

Write a function that counts the best exam twice and then takes the average. Use it to grade all of the students.

1. Write code that solves the problem for a real object  
2. Wrap the code in `function(){}` to save it  
3. Add the name of the real object as the function argument 

```{r}
vec <- exams[[1]]

grade <- function(vec) {
  (sum(vec) - min(vec)) / (length(vec) - 1)
}

grade(exams[[2]])

exams %>%
 map_dbl(grade)

grade <- function(x) {
(sum(x) - min(x)) / (length(x) - 1)
}
exams %>%
 map_dbl(grade)

exams %>%
map_dbl(function(x) (sum(x) - min(x)) / (length(x) - 1))

```

### Your Turn 8

Compute a final grade for each student, where the final grade is the average test score plus any `extra_credit` assigned to the student. Return the results as a double (i.e. numeric) vector.

```{r}

```


***

# Take Aways

Lists are a useful way to organize data, but you need to arrange manually for functions to iterate over the elements of a list.

You can do this with the `map()` family of functions in the purrr package.

To write a function, 

1. Write code that solves the problem for a real object  
2. Wrap the code in `function(){}` to save it  
3. Add the name of the real object as the function argument 

This sequence will help prevent bugs in your code (and reduce the time you spend correcting bugs). 

<!-- This file by Charlotte Wickham is licensed under a Creative Commons Attribution 4.0 International License, adapted from the orignal work at https://github.com/rstudio/master-the-tidyverse by RStudio. -->
