---
title: "Model"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
library(broom)

wages <- heights %>% filter(income > 0)
```

## Your Turn 1

Fit the model on the slide and then examine the output. What does it look like?

```{r}
mod_e <- lm(log(income) ~ education, data = wages)
mod_e

wages %>%
 ggplot(aes(log(income))) + geom_histogram(binwidth = 0.25)
```

## Your Turn 2

Use a pipe to model `log(income)` against `height`. Then use broom and dplyr functions to extract:

1. The **coefficient estimates** and their related statistics 
2. The **adj.r.squared** and **p.value** for the overall model

```{r, error = TRUE}
mod_e <-lm(log(income) ~ education, data = wages)
view(mod_e)
class(mod_e)
summary(mod_e) # give you all model parameters 

# can pass the data using a dot for data = .

mod_e <- wages %>%
 lm(log(income) ~ education, data = .)
summary(mod_e)

mod_e %>% tidy() # now this is a tibble and you can reach in and grab the estimates 

mod_e %>% glance()
mod_e %>% augment()
mod_e %>% augment(data = wages)

mod_h <- wages %>% 
  lm(log(income) ~ height, data = .)

estimates <- mod_h %>% tidy()
str(estimates)
fit <- mod_h %>% glance()
fit1 <- mod_h %>% glance() %>% select(adj.r.squared, p.value)
str(fit)
aug <- mod_h %>% augment()
str(aug)

```

## Your Turn 3

Model `log(income)` against `education` _and_ `height`. Do the coefficients change?

```{r, error = TRUE}
mod_eh <- wages %>% lm(log(income) ~ education + height, data = .)
summary(mod_eh)
mod_eh %>% 
  tidy()
wages %>% 
  ggplot() + geom_histogram(aes(x = education, binwidth = 0.25))
                            
```

## Your Turn 4

Model `log(income)` against `education` and `height` and `sex`. Can you interpret the coefficients?

```{r, error = TRUE}
mod_ehs <- wages %>%
 lm(log(income) ~ education + height + sex, data = .)
mod_ehs %>% 
  tidy()

```

## Your Turn 5

Use a broom function and ggplot2 to make a line graph of `height` vs `.fitted` for our heights model, `mod_h`.

_Bonus: Overlay the plot on the original data points._

```{r}
mod_h <- wages %>% lm(log(income) ~ height, data = .)
mod_h %>% tidy()
mod_h %>% glance()
height1 <- mod_h %>% augment()

mod_h %>%
 augment(data = wages) %>%
 ggplot(mapping = aes(x = height, y = .fitted)) +
 geom_point(mapping = aes(y = log(income)), alpha = 0.1) +
 geom_line(color = "blue")

```

## Your Turn 6

Repeat the process to make a line graph of `height` vs `.fitted` colored by `sex` for model `mod_ehs`. Are the results interpretable? Add `+ facet_wrap(~education)` to the end of your code. What happens?

```{r}
mod_ehs <- wages %>% lm(log(income) ~ education + height + sex, data = .)

mod_ehs %>%
 augment(data = wages) %>%
 ggplot(mapping = aes(x = height, y = .fitted, color = sex)) +
 geom_line() +
 facet_wrap(~ education)

```

## Your Turn 7

Use one of `spread_predictions()` or `gather_predictions()` to make a line graph of `height` vs `pred` colored by `model` for each of mod_h, mod_eh, and mod_ehs. Are the results interpretable? 

Add `+ facet_grid(sex ~ education)` to the end of your code. What happens?

```{r warning = FALSE, message = FALSE}
mod_h <- wages %>% lm(log(income) ~ height, data = .)
mod_eh <- wages %>% lm(log(income) ~ education + height, data = .)
mod_ehs <- wages %>% lm(log(income) ~ education + height + sex, data = .)

wages %>%
 spread_predictions(mod_h, mod_eh, mod_ehs)

wages %>%
 gather_predictions(mod_h, mod_eh, mod_ehs)

wages %>%
 gather_predictions(mod_h, mod_eh, mod_ehs) %>%
 ggplot(mapping = aes(x = height, y = pred, color = model)) +
 geom_line()

wages %>%
 gather_predictions(mod_h, mod_eh, mod_ehs) %>%
 ggplot(mapping = aes(x = height, y = pred, color = model)) +
 geom_line() +
 facet_grid(sex ~ education)

wages %>%
 gather_predictions(mod_h, mod_eh, mod_ehs) %>%
 filter(education > 11) %>%
 ggplot(mapping = aes(x = height, y = pred, color = sex)) +
 geom_line() +
 facet_grid(model ~ education)

```

***

# Take Aways

* Use `glance()`, `tidy()`, and `augment()` from the **broom** package to return model values in a data frame.

* Use `add_predictions()` or `gather_predictions()` or `spread_predictions()` from the **modelr** package to visualize predictions.

* Use `add_residuals()` or `gather_residuals()` or `spread_residuals()` from the **modelr** package to visualize residuals.

<!-- This file by Charlotte Wickham is licensed under a Creative Commons Attribution 4.0 International License, adapted from the orignal work at https://github.com/rstudio/master-the-tidyverse by RStudio. -->

